---
title: Elastic Net cross-validation demo
author: Peter Carbonetto
output: workflowr::wflow_html
---

TO DO: Give brief overview of the this demo.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load a couple packaeges and the cross-validation code.

```{r load-pkgs, message=FALSE}
library(glmnet)
library(parallel)
source("../code/cv.R")
```

Initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

Simulate a regression data set.

```{r sim-data}
n <- 200
p <- 1000
X <- matrix(rnorm(n*p),n,p)
b <- rep(0,p)
b[sample(p,4)] <- c(1,-1,1,-1)
y <- X %*% b + rnorm(n)
```

To perform cross-validation, we need to define three functions: (1) a
function to fit an Elastic Net model; and (2) a function to predict Y
from X and a fitted model; and (3) a function to evaluate the accuracy
of the predicted Y (here we use the mean-squared error, which is also
what is used in the glmnet package).

This function fits an Elastic Net model using glmnet:

```{r fit-glmnet}
fit_glmnet <- function (x, y, cvpar, noncvpar, init)
  glmnet(x,y,lambda = cvpar,alpha = 0.5)
  ```
  
This function preducts Y using the fitted Elastic Net model:

```{r predict-glmnet}
predict_glmnet <- function (x, model)
  predict(model,x)
```

This function computes the mean squared error (MSE) between the
predicted Y and true Y:

```{r compute-mse}
compute_mse <- function (pred, true)
  mean((pred - true)^2)
```

Having defined these three functions, we are ready to use the
`perform_cv` interface to run the cross-validation to choose the
strength ($\lambda$) of the Elastic Net penalty:

```{r run-cv}
lambda <- round(rev(exp(seq(-3.75,0.85,length.out = 100))),digits = 4)
t0 <- proc.time()
cv <- perform_cv(fit_glmnet,predict_glmnet,compute_mse,X,y,lambda,nc = 2)
t1 <- proc.time()
print(t1 - t0)
```

Compare with cv.glmnet:

```{r compare-to-glmnet, fig.height=4, fig.width=4}
res <- cv.glmnet(X,y,alpha = 0.5)
plot(res$lambda,res$cvm,type = "l",col = "darkblue",lwd = 2,log = "x",
     xlab = "lambda",ylab = "mse")
lines(lambda,rowMeans(cv),col = "darkorange",lwd = 2,lty = "dashed")
```
