---
title: Elastic Net cross-validation demo
author: Peter Carbonetto
output: workflowr::wflow_html
---

Here we illustrate how to use the `perform_cv` cross-validation
interface to estimate the penalty strength paramter in the Elastic Net
model in glmnet. Of course, glmnet already has a cross-validation
interface, so we can use the existing cross-validation function in
glmnet to verify our implementation. Indeed, `perform_cv` almost
exactly reproduces the `cv.glmnet` output in this example.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center",dpi = 120)
```

Load a couple packages and the `perform_cv` code.

```{r load-pkgs, message=FALSE}
library(glmnet)
library(parallel)
source("../code/cv.R")
```

Initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

Simulate a regression data set.

```{r sim-data}
n <- 200
p <- 1000
X <- matrix(rnorm(n*p),n,p)
b <- rep(0,p)
b[sample(p,4)] <- c(1,-1,1,-1)
y <- X %*% b + rnorm(n)
```

To perform cross-validation, we need to define three functions: (1) a
function to fit an Elastic Net model; (2) a function to predict Y
using the fitted Elastic Net model; and (3) a function to evaluate the
accuracy of the predicted Y (here we use the mean-squared error, which
is also what is used in the glmnet package to evaluate the predictions).

This function fits an Elastic Net model:

```{r fit-glmnet}
fit_glmnet <- function (x, y, cvpar, noncvpar, init)
  glmnet(x,y,lambda = cvpar,alpha = 0.5)
  ```
  
This function predicts Y using the fitted Elastic Net model:

```{r predict-glmnet}
predict_glmnet <- function (x, model)
  predict(model,x)
```

This function computes the mean squared error (MSE) between the
predicted and true Y:

```{r compute-mse}
compute_mse <- function (pred, true)
  mean((pred - true)^2)
```

Having defined these three functions, we are ready to use
`perform_cv`:

```{r run-cv}
lambda <- round(rev(exp(seq(-3.75,0.85,length.out = 100))),digits = 4)
t0 <- proc.time()
cv <- perform_cv(fit_glmnet,predict_glmnet,compute_mse,X,y,lambda,nc = 2)
t1 <- proc.time()
print(t1 - t0)
```

Compare the result with the result obtained from running `cv.glmnet`
on the same data (the dark blue line is the `cv.glmnet` result, and
the dashed, orange line is our result):

```{r compare-to-glmnet, fig.height=2.8, fig.width=3}
par(mar = c(4,4,0,0))
res <- cv.glmnet(X,y,alpha = 0.5)
plot(res$lambda,res$cvm,type = "l",col = "darkblue",lwd = 2,log = "x",
     xlab = "lambda",ylab = "mse")
lines(lambda,rowMeans(cv),col = "darkorange",lwd = 2,lty = "dashed")
```
